Aerodigestive sampling reveals altered microbial exchange between lung, oropharyngeal, and gastric microbiomes in children with impaired swallow function

### 2018-12-10 aspiration/plosone_aspiration_main.tex
Add refs to intro, fix a few cosmetic things esp in TOC
Finalize aspiration paper + supplement

\section{Abstract}

\subsection{Background}

@@

\subsection{Methods}

@@

\subsection{Results}

@@

\subsection{Conclusions}

@@

\subsection{Key words}

Aerodigestive microbiomes; respiratory, oral, and gastric microbiomes; gastroesophageal reflux; impedance; aspiration; video fluoroscopic swallow study

\section{Introduction}

@@

Previous studies have shown that the mouth, upper respiratory tract, and lung microbiota contain similar microbes, and that upstream oral communities seed downstream sites (e.g. lungs and stomach) \cite{Bassis2015source,Charlson2011topographical,rosen2015ppi}.
However, there is little consensus on whether there exists a distinct or "core" lung microbiome that is consistent across people \cite{Charlson2011topographical,venkataraman-2015-dispersal,segal-2013-pneumotypes,erbDownward-2011-COPD}.
Most studies, however, agree that the lung microbial communities share taxa with the oral microbiome, but that there are some bacteria present in lung communities whose abundances cannot be traced solely to the mouth \cite{Charlson2011topographical,Bassis2015source,morris-2013-healthsmokers,segal-2013-pneumotypes}.

@@

We first performed brushing of the posterior tongue to obtain oropharyngeal samples, placing the brush in TE buffer at -80˚C.
Second, the bronchoscopy and bronchoalveolar lavage (BAL) was performed through an endotracheal tube in distal airways of the right middle lung or the most visually inflamed lung.
Finally, gastric sampling was performed during the EGD.
The endoscope was advanced, without suctioning, immediately into the stomach where the gastric fluid was suctioned into a sterile leukitrap.
A minimum of 1 cc of gastric and lung fluid were collected and transferred to -80˚C.
Each patient had a triad of samples collected: oropharynx, gastric fluid, and BAL (Tables \ref{tab1} and \ref{tab2}) \cite{rosen2015ppi}.

@@

\subsection{Availability of data and materials}
Code to reproduce the analyses presented here are available at www.github.com/cduvallet/aspiration-analysis-public.
The 16S sequencing data used in this study are available in the SRA repository at accession number SRP141148 and clinical metadata are available upon request from the corresponding author.

@@

Two hundred and twenty two patients were included in the analysis (Tables ref{tab1} and \ref{tab2}).
The mean age of the patients was 7.1 $\pm$ 5.4 years.
One hundred out of 222 patients were taking proton pump inhibitors at the time of sampling.
One hundred and four patients had a videoflouroscopic swallow study of which 47 (45\%) had evidence of aspiration or penetration and 57 (55\%) had normal swallow function.
Thirty one patients had pH-MII testing for gastroesophageal reflux at the time of sample collection.

@@

To identify specific microbes exchanging between sites, we reasoned that an actively exchanging microbe's abundances in two sites should be correlated across patients (Supplementary Figure \ref{figs1} and Methods).
We identified 12 OTUs exchanged between lung and oropharyngeal, 74 between gastric fluid and lung, and 118 between oropharyngeal and gastric fluid communities.
These results were statistically significant: we found a maximum of 2 exchanged OTUs between sites in our null analysis.
The low number of directly exchanged OTUs between the oropharynx and lungs supports the finding that these sites are more distinct than others in the aerodigestive tract.
The lungs and stomach exchange fewer OTUs than the oropharynx and stomach even though they have comparable intra-patient similarities, suggesting that factors other than specific bacterial exchange contributes to the similarity between lungs and stomachs within patients.

@@

To identify potential microbial biomarkers of aspiration, we looked at the exchanged OTUs which were most frequently present in the lung and oropharyngeal communities of aspirators relative to non-aspirators.
In the oropharyngeal-lung exchanged OTUs, these were an unknown OTU in the \textit{Flavobacteriaceae} family, OTUs in the \textit{Fusobacterium}, \textit{Rothia}, \textit{Veillonella} genera, and an unknown OTU in the \textit{Prevotellaceae} family, among others (Table \ref{tab4}, gastric-lung OTUs in Supplementary Table \ref{tabs1}).

We used Random Forest classifiers trained on the presence of exchanged OTUs in different sites to test their potential as biomarkers.
The concordant presence or absence of exchanged OTUs in the two sites improved classifiers based on the oropharyngeal-lung OTUs but not the ones based on the lung-gastric OTUs, relative to classifiers based on the presence of the exchanged OTUs in either site alone (Table \ref{tab5}).
Using Random Forest classifiers trained on the entire microbiomes, we found that combining the oropharynx and lung communities resulted in a better classifier than either community alone (Table \ref{tab6}).
Surprisingly, the classifiers trained on oropharyngeal and gastric communities performed well, despite our expectation that aspiration-induced changes in the microbiome would manifest in the lungs rather than the oropharynx or stomach.
We confirmed that the patients' aspiration status was not confounded with proton pump inhibitor usage (Fisher exact p-value $>$ 0.2), but there may be other co-morbidities or unmeasured confounders that could be driving the differences detected in these communities.
However, taken together, these results suggest that identifying a biomarker for aspiration based on bacteria in both the lungs and oropharynx may be possible, and that these two sites together contain more information about a patient’s aspiration status than either site alone.

@@

There are several limitations to our study.
First, because it is unethical to perform bronchoscopies on healthy children, our patients in this study had respiratory symptoms.
However, we believe that our patient population represents patients typically seen in aerodigestive centers and that understanding the degree of microbial exchange is most clinically relevant in patients with symptoms.
The microbial populations we found in this study are similar to those of previously published studies of both healthy and symptomatic adults which reinforces the validity of our results \cite{Bassis2015source,Charlson2011topographical,erbDownward-2011-COPD,morris-2013-healthsmokers}.
Second, the number of patients undergoing pH-MII testing was relatively small which limits our conclusions about the impact of gastroesophageal reflux on the lung.
However, our study raises enough concerns about the significance of oropharyngeal-lung exchange in children with impaired swallowing that gastroesophageal reflux should not be considered as the primary source of microbial exchange causing pulmonary symptoms.
Third, the diagnosis of oropharyngeal dysphagia in this study was based on VFSS.
While this only categorizes patients based on a “one-point-in-time” study, it is the gold standard test to diagnose oropharyngeal dysphagia in children and therefore we feel it is appropriate for use in this study.

@@

\begin{table}
\begin{center}
\begin{tabular}{ccccc}
 & Normal & Aspirators & Not tested & Total \\
 \midrule
BAL & 33 & 33 & 36 & 102 \\
Oropharyngeal swab & 43 & 36 & 97 & 176 \\
Gastric fluid & 48 & 41 & 58 & 147 \\
Stool & & & 20 & 20 \\
\bottomrule
\end{tabular}
\caption{Number of patient samples for each body site.}\label{tab1}
\end{center}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{ccccc}
 & Normal & Aspirators & Not tested & Total \\
 \midrule
BAL and oropharyngeal swab & 23 & 25 & 25 & 73 \\
BAL and gastric fluid & 28 & 29 & 32 & 89 \\
Gastric fluid and oropharyngeal swab & 35 & 32 & 45 & 112 \\
Stool and oropharyngeal swab & & & 20 & 20 \\
\bottomrule
\end{tabular}
\caption{Number of patients with multiple body sites sequenced.}\label{tab2}
\end{center}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{cccc}
	Within people & & Across people & p \\
	\midrule
	Lung and oropharynx & more different & oropharynx & $< 1 \times 10^{-8}$ \\
	& not significant & lungs & 0.8 \\
	\midrule
	Lung and gastric fluid & more similar & lungs & $< 1 \times 10^{-11}$ \\
	& more similar & gastric & $< 1 \times 10^{-8}$ \\
	\midrule
	Gastric and oropharyngeal & more similar & gastric & $< 1 \times 10^{-11}$ \\
	& not significant & oropharyngeal & 0.07 \\
	\bottomrule
\end{tabular}
\caption{\textbf{Lung and gastric microbial communities are driven primarily by person rather than body site.} For each patient and each aerodigestive site, we compared the average JSD between that patient's site and all other patients' communities of that same site with the JSD between that patient's site and their other two aerodigestive sites. For example, the top row shows the comparisons between (1) the average JSD between a patient's oropharyngeal community and all other oropharyngeal communities and (2) the JSD between that patient's own oropharyngeal and lung communities. We subtracted each patient's between-sites JSD from their average between-patient JSD and calculated Wilcoxon signed-rank p-values using Python's \texttt{scipy.stats.wilcoxon} function.}\label{tab3}
\end{center}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{ccccc}
	Family & Genus & Non-aspirator & Aspirator & Difference \\
	\midrule
	Flavobacteriaceae &  & 8.7 & 48.0 & 39.3 \\
	Fusobacteriaceae & Fusobacterium & 30.4 & 68.0 & 37.6 \\
	Micrococcaceae & Rothia & 8.7 & 44.0 & 35.3 \\
	Veillonellaceae & Veillonella & 26.1 & 60.0 & 33.9 \\
	Prevotellaceae &  & 43.5 & 76.0 & 32.5 \\
	Porphyromonadaceae & Porphyromonas & 39.1 & 68.0 & 28.9 \\
	Streptococcaceae & Streptococcus & 13.0 & 40.0 & 27.0 \\
	Veillonellaceae & Centipeda & 8.7 & 32.0 & 23.3 \\
	Prevotellaceae & Prevotella & 17.4 & 36.0 & 18.6 \\
	Leptotrichiaceae & Streptobacillus & 21.7 & 40.0 & 18.3 \\
	Fusobacteriaceae & Fusobacterium & 17.4 & 32.0 & 14.6 \\
	Aerococcaceae & Abiotrophia & 21.7 & 28.0 & 6.3 \\
	\bottomrule
\end{tabular}
\caption{\textbf{Prevalence of lung-oropharynx exchanged OTUs.} Prevalence is calculated as the percentage of patients who have the OTU present in both their lungs and oropharynx, calculated separately among aspirators (N $=$ 25) and non-aspirators (N $=$ 23). OTUs are ordered by their differential prevalence in aspirators relative to non-aspirators, and are labeled with their family- and genus-level taxonomies. Blank genus names indicate OTUs which were not annotated at the genus level. A similar table for the lung-gastric exchange OTUs can be found in Supplementary Table 1.}\label{tab4}
\end{center}
\end{table}

%% Exchange classifiers based on presence
% These values are copied manually from 2018-01-19.aspiration_classifiers_results_final
\begin{table}
\begin{center}
%\begin{tabular}{ll}

@@

\caption{\textbf{Classifiers based on the presence of exchanged OTUs.} (Top) Classifiers built from the presence of lung-oropharynx exchanged OTUs. (Bottom) Classifiers built from the presence of lung-gastric exchanged OTUs. Rows indicate which microbial community was used to train each classifier. In the ``concordance'' classifiers, OTUs which were either present or absent in both sites were coded as 1 and OTUs which were present in one site but absent in the other were coded as 0. AUCs are calculated as the area under the average ROC curve from five-fold cross validation. Fisher's exact p values are calculated on the predictions on the hold-out data for all cross validation folds. Each classifier was built 100 times with random patient splits and classifier initializations, and mean values are reported here. Similar classifiers built from the abundance of exchanged OTUs are shown in Supplementary Table 2. AUCs and Fisher p-values from all 100 repetitions for all classifiers are shown in Supplementary Figures \ref{asp_figs2} and \ref{asp_figs3}.) }\label{tab5}
\end{center}
\end{table}

% These have been updated from 2018-01-17.aspiration_classifiers_full_community_final
\begin{table}
\begin{center}
\begin{tabular}{cccc}
  Sites & AUC & Fisher p-value & N (non-asp/asp) \\
 \midrule
Lung & 0.66	& 0.2 & 33/33 \\
Oropharyngeal swab & 0.71 & 0.02 & 43/36 \\
Gastric fluid & 0.67 & 0.11 & 48/41 \\
Lung and oropharyngeal swab & 0.81 & 0.01 & 23/25 \\
Lung and gastric fluid & 0.70 & 0.07 & 28/29 \\
Oropharyngeal swab and gastric fluid & 0.76 & 0.02 & 35/32 \\
All three sites & 0.83 & 0.01 & 19/23 \\
\bottomrule
\end{tabular}
\caption{\textbf{Classifiers based on perturbed relationship between lung and oropharyngeal microbiota can distinguish aspirators from non-aspirators.} Areas under the ROC curve (AUC) and Fisher p-values calculated from classifiers trained on the entire microbial communities. Each row is a different classifier based on different combinations of aerodigestive communities, indicated in the ``Sites'' column. In the multi-site classifiers, the abundances of OTUs in different sites were used as separate features. For each classifier type, 100 classifiers were built, with random patient splits and classifier initializations. Mean values are reported. The distribution of AUCs and Fisher p-values from all 100 repetitions are shown in Supplementary Figure \ref{figs4}.}\label{tab6}
\end{center}
\end{table}

% This table is from Rachel
\begin{table}
\begin{center}
\begin{tabular}{cc}
 & Mean (std) \\
 \midrule
Number of acid episodes & 24.1 (26.9) \\
Number of nonacid episodes & 14.8 (16.4) \\
Number of pH only episodes & 16.6 (14.8) \\
Number of total reflux episodes & 38.9 (33.8) \\
Percent time proximal reflux & 0.005 (0.005) \\
Percent time distal reflux & 0.012 (0.011) \\
Percent time pH $<$ 4 & 7.2 (11.5) \\
Number abnormal by pH-metry & 9 \\
Number abnormal by MII & 4 \\
\bottomrule
\end{tabular}
\caption{Reflux characteristics measured by pH-MII.}\label{tab7}
\end{center}
\end{table}

>>>>

\section*{Abstract}

\subsection*{Background}

@@

\subsection*{Methods}

@@

\subsection*{Results}

@@

\subsection*{Conclusions}

@@

\newpage
\section{Introduction}

@@

Previous studies have shown that the mouth, upper respiratory tract, and lung microbiota contain similar microbes, and that upstream oral communities seed downstream sites (e.g. lungs and stomach) \cite{Bassis2015source,Charlson2011topographical,rosen2015ppi}.
However, there is little consensus on whether there exists a distinct or "core" lung microbiome that is consistent across people \cite{Charlson2011topographical,venkataraman-2015-dispersal,segal-2013-pneumotypes,erbDownward-2011-COPD}.
Most studies, however, agree that the lung microbial communities share taxa with the oral microbiome, but that there are some bacteria present in lung communities whose abundances cannot be traced solely to the mouth \cite{Bassis2015source,Charlson2011topographical,morris-2013-healthsmokers,segal-2013-pneumotypes}.

@@

We first performed brushing of the posterior tongue to obtain oropharyngeal samples, placing the brush in TE buffer at -80C.
Second, the bronchoscopy and bronchoalveolar lavage (BAL) was performed through an endotracheal tube in distal airways of the right middle lung or the most visually inflamed lung.
Finally, gastric sampling was performed during the EGD.
The endoscope was advanced, without suctioning, immediately into the stomach where the gastric fluid was suctioned into a sterile leukitrap.
A minimum of 1 cc of gastric and lung fluid were collected and transferred to -80C.
Each patient had a triad of samples collected: oropharynx, gastric fluid, and BAL (Tables \ref{tab1} and \ref{tab2}) \cite{rosen2015ppi}.

@@

\subsection{Availability of data and materials}
Code to reproduce the analyses presented here are available at www.github.com/cduvallet/aspiration-analysis-public.
The 16S sequencing data used in this study will be made available in the SRA repository at accession number SRP141148 upon publication and clinical metadata are available upon request from the corresponding author.

@@

Two hundred and twenty two patients were included in the analysis (Tables \ref{tab1} and \ref{tab2}).
The mean age of the patients was 7.1 $\pm$ 5.4 years.
One hundred out of 222 patients were taking proton pump inhibitors at the time of sampling.
One hundred and four patients had a videoflouroscopic swallow study of which 47 (45\%) had evidence of aspiration or penetration and 57 (55\%) had normal swallow function.
Thirty one patients had pH-MII testing for gastroesophageal reflux at the time of sample collection.

@@

To identify specific microbes exchanging between sites, we reasoned that an actively exchanging microbe's abundances in two sites should be correlated across patients (Supplementary Figure A-1 and Methods).
We identified 12 OTUs exchanged between lung and oropharyngeal, 74 between gastric fluid and lung, and 118 between oropharyngeal and gastric fluid communities.
These results were statistically significant: we found a maximum of 2 exchanged OTUs between sites in our null analysis.
The low number of directly exchanged OTUs between the oropharynx and lungs supports the finding that these sites are more distinct than others in the aerodigestive tract.
The lungs and stomach exchange fewer OTUs than the oropharynx and stomach even though they have comparable intra-patient similarities, suggesting that factors other than specific bacterial exchange contributes to the similarity between lungs and stomachs within patients.

@@

To identify potential microbial biomarkers of aspiration, we looked at the exchanged OTUs which were most frequently present in the lung and oropharyngeal communities of aspirators relative to non-aspirators.
In the oropharyngeal-lung exchanged OTUs, these were an unknown OTU in the \textit{Flavobacteriaceae} family, OTUs in the \textit{Fusobacterium}, \textit{Rothia}, \textit{Veillonella} genera, and an unknown OTU in the \textit{Prevotellaceae} family, among others (Table \ref{tab4}, gastric-lung OTUs in Supplementary Table A-1).

We used Random Forest classifiers trained on the presence of exchanged OTUs in different sites to test their potential as biomarkers.
The concordant presence or absence of exchanged OTUs in the two sites improved classifiers based on the oropharyngeal-lung OTUs but not the ones based on the lung-gastric OTUs, relative to classifiers based on the presence of the exchanged OTUs in either site alone (Table \ref{tab5}).
Using Random Forest classifiers trained on the entire microbiomes, we found that combining the oropharynx and lung communities resulted in a better classifier than either community alone (Table \ref{tab6}).
Surprisingly, the classifiers trained on oropharyngeal and gastric communities performed well, despite our expectation that aspiration-induced changes in the microbiome would manifest in the lungs rather than the oropharynx or stomach.
We confirmed that the patients' aspiration status was not confounded with proton pump inhibitor usage (Fisher exact p-value $>$ 0.2), but there may be other co-morbidities or unmeasured confounders that could be driving the differences detected in these communities.
However, taken together, these results suggest that identifying a biomarker for aspiration based on bacteria in both the lungs and oropharynx may be possible, and that these two sites together contain more information about a patient's aspiration status than either site alone.

@@

There are several limitations to our study.
First, because it is unethical to perform bronchoscopies on healthy children, our patients in this study had respiratory symptoms.
However, we believe that our patient population represents patients typically seen in aerodigestive centers and that understanding the degree of microbial exchange is most clinically relevant in patients with symptoms.
The microbial populations we found in this study are similar to those of previously published studies of both healthy and symptomatic adults which reinforces the validity of our results \cite{Bassis2015source,Charlson2011topographical,erbDownward-2011-COPD,morris-2013-healthsmokers}.
Second, the number of patients undergoing pH-MII testing was relatively small which limits our conclusions about the impact of gastroesophageal reflux on the lung.
However, our study raises enough concerns about the significance of oropharyngeal-lung exchange in children with impaired swallowing that gastroesophageal reflux should not be considered as the primary source of microbial exchange causing pulmonary symptoms.
Third, the diagnosis of oropharyngeal dysphagia in this study was based on VFSS.
While this only categorizes patients based on a ``one-point-in-time'' study, it is the gold standard test to diagnose oropharyngeal dysphagia in children and therefore we feel it is appropriate for use in this study.

@@

\begin{table}[H]
\begin{center}
\begin{tabular}{ccccc}
 & Normal & Aspirators & Not tested & Total \\
 \midrule
BAL & 33 & 33 & 36 & 102 \\
Oropharyngeal swab & 43 & 36 & 97 & 176 \\
Gastric fluid & 48 & 41 & 58 & 147 \\
Stool & & & 20 & 20 \\
\bottomrule
\end{tabular}
\caption{Number of patient samples for each body site.}\label{tab1}
\end{center}
\end{table}

\begin{table}[H]
\begin{center}
\begin{tabular}{ccccc}
 & Normal & Aspirators & Not tested & Total \\
 \midrule
BAL and oropharyngeal swab & 23 & 25 & 25 & 73 \\
BAL and gastric fluid & 28 & 29 & 32 & 89 \\
Gastric fluid and oropharyngeal swab & 35 & 32 & 45 & 112 \\
Stool and oropharyngeal swab & & & 20 & 20 \\
\bottomrule
\end{tabular}
\caption{Number of patients with multiple body sites sequenced.}\label{tab2}
\end{center}
\end{table}

\begin{table}[H]
\begin{center}
\begin{tabular}{cccc}
	Within people & & Across people & p \\
	\midrule
	Lung and oropharynx & more different & oropharynx & $< 1 \times 10^{-8}$ \\
	& not significant & lungs & 0.8 \\
	\midrule
	Lung and gastric fluid & more similar & lungs & $< 1 \times 10^{-11}$ \\
	& more similar & gastric & $< 1 \times 10^{-8}$ \\
	\midrule
	Gastric and oropharyngeal & more similar & gastric & $< 1 \times 10^{-11}$ \\
	& not significant & oropharyngeal & 0.07 \\
	\bottomrule
\end{tabular}
\caption{\textbf{Lung and gastric microbial communities are driven primarily by person rather than body site.} For each patient and each aerodigestive site, we compared the average JSD between that patient's site and all other patients' communities of that same site with the JSD between that patient's site and their other two aerodigestive sites. For example, the top row shows the comparisons between (1) the average JSD between a patient's oropharyngeal community and all other oropharyngeal communities and (2) the JSD between that patient's own oropharyngeal and lung communities. We subtracted each patient's between-sites JSD from their average between-patient JSD and calculated Wilcoxon signed-rank p-values using Python's \texttt{scipy.stats.wilcoxon} function.}\label{tab3}
\end{center}
\end{table}

\begin{table}[H]
\begin{center}
\begin{tabular}{ccccc}
	Family & Genus & Non-aspirator & Aspirator & Difference \\
	\midrule
	Flavobacteriaceae &  & 8.7 & 48.0 & 39.3 \\
	Fusobacteriaceae & Fusobacterium & 30.4 & 68.0 & 37.6 \\
	Micrococcaceae & Rothia & 8.7 & 44.0 & 35.3 \\
	Veillonellaceae & Veillonella & 26.1 & 60.0 & 33.9 \\
	Prevotellaceae &  & 43.5 & 76.0 & 32.5 \\
	Porphyromonadaceae & Porphyromonas & 39.1 & 68.0 & 28.9 \\
	Streptococcaceae & Streptococcus & 13.0 & 40.0 & 27.0 \\
	Veillonellaceae & Centipeda & 8.7 & 32.0 & 23.3 \\
	Prevotellaceae & Prevotella & 17.4 & 36.0 & 18.6 \\
	Leptotrichiaceae & Streptobacillus & 21.7 & 40.0 & 18.3 \\
	Fusobacteriaceae & Fusobacterium & 17.4 & 32.0 & 14.6 \\
	Aerococcaceae & Abiotrophia & 21.7 & 28.0 & 6.3 \\
	\bottomrule
\end{tabular}
\caption{\textbf{Prevalence of lung-oropharynx exchanged OTUs.} Prevalence is calculated as the percentage of patients who have the OTU present in both their lungs and oropharynx, calculated separately among aspirators (N $=$ 25) and non-aspirators (N $=$ 23). OTUs are ordered by their differential prevalence in aspirators relative to non-aspirators, and are labeled with their family- and genus-level taxonomies. Blank genus names indicate OTUs which were not annotated at the genus level. A similar table for the lung-gastric exchange OTUs can be found in Supplementary Table A-1.}\label{tab4}
\end{center}
\end{table}

%% Exchange classifiers based on presence
% These values are copied manually from 2018-01-19.aspiration_classifiers_results_final
\begin{table}[H]
\begin{center}
%\begin{tabular}{ll}

@@

\caption{\textbf{Classifiers based on the presence of exchanged OTUs.} (Top) Classifiers built from the presence of lung-oropharynx exchanged OTUs. (Bottom) Classifiers built from the presence of lung-gastric exchanged OTUs. Rows indicate which microbial community was used to train each classifier. In the ``concordance'' classifiers, OTUs which were either present or absent in both sites were coded as 1 and OTUs which were present in one site but absent in the other were coded as 0. AUCs are calculated as the area under the average ROC curve from five-fold cross validation. Fisher's exact p values are calculated on the predictions on the hold-out data for all cross validation folds. Each classifier was built 100 times with random patient splits and classifier initializations, and mean values are reported here. Similar classifiers built from the abundance of exchanged OTUs are shown in Supplementary Table A-2. AUCs and Fisher p-values from all 100 repetitions for all classifiers are shown in Supplementary Figures A-2 and A-3.) }\label{tab5}
\end{center}
\end{table}

% These have been updated from 2018-01-17.aspiration_classifiers_full_community_final
\begin{table}[H]
\begin{center}
\begin{tabular}{cccc}
  Sites & AUC & Fisher p-value & N (non-asp/asp) \\
 \midrule
Lung & 0.66	& 0.2 & 33/33 \\
Oropharyngeal swab & 0.71 & 0.02 & 43/36 \\
Gastric fluid & 0.67 & 0.11 & 48/41 \\
Lung and oropharyngeal swab & 0.81 & 0.01 & 23/25 \\
Lung and gastric fluid & 0.70 & 0.07 & 28/29 \\
Oropharyngeal swab and gastric fluid & 0.76 & 0.02 & 35/32 \\
All three sites & 0.83 & 0.01 & 19/23 \\
\bottomrule
\end{tabular}
\caption{\textbf{Classifiers based on perturbed relationship between lung and oropharyngeal microbiota can distinguish aspirators from non-aspirators.} Areas under the ROC curve (AUC) and Fisher p-values calculated from classifiers trained on the entire microbial communities. Each row is a different classifier based on different combinations of aerodigestive communities, indicated in the ``Sites'' column. In the multi-site classifiers, the abundances of OTUs in different sites were used as separate features. For each classifier type, 100 classifiers were built, with random patient splits and classifier initializations. Mean values are reported. The distribution of AUCs and Fisher p-values from all 100 repetitions are shown in Supplementary Figure A-4.}\label{tab6}
\end{center}
\end{table}

% This table is from Rachel
\begin{table}[H]
\begin{center}
\begin{tabular}{cc}
 & Mean (std) \\
 \midrule
Number of acid episodes & 24.1 (26.9) \\
Number of nonacid episodes & 14.8 (16.4) \\
Number of pH only episodes & 16.6 (14.8) \\
Number of total reflux episodes & 38.9 (33.8) \\
Percent time proximal reflux & 0.005 (0.005) \\
Percent time distal reflux & 0.012 (0.011) \\
Percent time pH $<$ 4 & 7.2 (11.5) \\
Number abnormal by pH-metry & 9 \\
Number abnormal by MII & 4 \\
\bottomrule
\end{tabular}
\caption{Reflux characteristics measured by pH-MII.}\label{tab7}
\end{center}
\end{table}
